\documentclass[12pt, a4paper, roman]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')

% modern themes
\moderncvstyle{banking}
\moderncvcolor{blue}              
\nopagenumbers{}                                  % uncomment to suppress automatic page numbering for CVs longer than one page

% character encoding
\usepackage[utf8]{inputenc}
\usepackage{fontawesome}
\usepackage{tabularx}
\usepackage{ragged2e}

% adjust the page margins
\usepackage[scale=0.8]{geometry}
\usepackage{multicol}

\usepackage{import}
\setlength{\tabcolsep}{12pt}
% personal data
\name{Pratyush}{Kumar}

%----------------------------------------------------------------------------------
%            content
%----------------------------------------------------------------------------------
\begin{document}

\makecvtitle
\vspace*{-15mm}
\centering{PhD researcher --- Control Systems and Machine Learning}

\justifying
\vspace{0.1in}
To, \\
The Platform Architecture Group, Apple \hfill November 29, 2020 \\

\vspace{0.1in}
My name is Pratyush Kumar, I am a fourth-year PhD candidate at the University of California, Santa Barbara (UCSB). I am writing to express my interest in the \textbf{Controls Engineering Internship} position at Apple for summer 2021.

\vspace{0.1in}
To give the recruitment committee an idea about my background - for my 
PhD research, I am working on 
the uses of machine learning 
for control and optimization of 
large-scale chemical processes 
and commercial heating, ventilation, and air-conditioning 
(HVAC) buildings. I have acquired knowledge and expertise in \textbf{applied machine learning and control systems theory/algorithms}.
I am versed in control and state estimation technologies such as \textbf{model predictive control, linear quadratic regulator, Kalman filtering, moving horizon estimation}, etc. I also have experience in system identification techniques via PhD research projects and a
previous internship at Johnson Controls, 
where I worked on developing temperature dynamic models for use in 
real-time energy optimization at commercial HVAC sites.
I have one full-length manuscript under review in a chemical 
engineering journal and more to be submitted related to control and 
system identification topics.

\vspace{0.1in}
I am interested in the application of controls, optimization, and machine learning techniques for a wide range of dynamic systems.
I am intrigued by the work done in the Platform Architecture Group 
at Apple, especially with the applications of control algorithms 
for thermal control and management.
I have gained the following relevant experiences via 
PhD research, previous internship, and
open-source software development,
which I believe can help me contribute fruitfully 
as a controls engineering intern at Apple:

\vspace{0.1in}
\begin{itemize}
	\item During my research and internship, 
	I have developed an in-depth understanding of 
	control and system identification algorithms relevant 
	for deployment in industrial applications.
	\vspace{0.15in}
	\item As an \textbf{open-source software developer and maintainer},
	I have gained experience in identifying software needs 
	of a community, contribute maintainable and sustainable 
	code for accelerating research and development. I am 
	proficient in the programing languages \textbf{C++} and \textbf{Python}.
\end{itemize}

\vspace{0.1in}
I am encouraged by the scale of impact possible at a company such as Apple, 
especially as a control systems engineer.
I am deeply committed to making progress in the internship using 
my control knowledge and programming skills.

\vspace{0.1in}
I thank the internship 
recruitment team for reviewing my application. I would love 
an opportunity to interview with the Platform Architecture Group in 
a video call to talk more about my background and experiences.

\vspace{0.1in}
Yours Sincerely, \\
Pratyush Kumar

\vspace{0.1in}
Attached: curriculum vit\ae{}
%\makeletterclosing

\end{document}
